---
title: "LRST_tutorial"
output: md_document
# rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LRST_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# LRST: Longitudinal Rank-Sum Test for Multivariate Longitudinal Data

The **LRST** R package implements the Longitudinal Rank-Sum Test, a nonparametric, rank-based omnibus test statistic designed for comprehensive assessment of treatment efficacy across multiple longitudinal primary endpoints in clinical trials. It effectively controls Type I error while enhancing statistical power, offering flexibility against various data distributions encountered in clinical research.


## Key Features

- **Comprehensive Assessment**: Evaluates treatment effects across multiple endpoints and time points without requiring multiplicity adjustments.
- **Enhanced Statistical Power**: Effectively controls Type I error, reducing the need for larger sample sizes.
- **Flexibility**: Robust against various data distributions, making it suitable for complex clinical trial data.
- **Efficient Data Utilization**: Maximizes the use of longitudinal data, capturing dynamic changes across multiple endpoints.



# Installation

You can install the package using:
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("devtools")
devtools::install_github("djghosh1123/LRST")
library(LRST)
```

# Example

The `lrst.2arm` function perform LRST on placebo and treatment data, while the `lrst.MultiArm` perform LRST on multi-arm clinical trials. Before perfomring LRST let us see what the data should look like.

# Simulated Data Generation


The below code generates the mean and standard deviation curves for the plaecbo group. The numbers are taken from the BAPI 302 trial.
```{r placebo}
xm = -t(cbind(c(0, 0.738, 1.313, 3.109, 4.525, 5.864, 7.338),
              c(0, 0.668, 3.975, 5.931, 8.259, 11.989, 13.976)))

xs = t(cbind(c(0, 4.79, 5.43, 6.54, 7.37, 8.15, 9.11),
             c(0, 10.27, 12.85, 14.95, 15.35, 16.87, 18.19)))
```
Now let's provide the dimentions for the simulation.

```{r Dimension}
N = 500 # Total Number of Patients
nx = 2/5*N # Niumber of patients in Placebo
ny = 3/5*N # Number of patients in Treatment
K = 2 # Number of Outcomes
T = 6 # Number of visits other than baseline
```

Let's define the temporal covariance structure for the two outcomes.Suppose the correlation between the outcomes is 0.5.


```{r Covariance}
sigma_u <- c(3.477, 7.232) # Random Effect Standard Deviation for Outcomes 
corr <- matrix(c(1,0.5, 0.5, 1),2,2)  # Correlation between the outcomes

sigma = diag(sigma_u) %*% corr %*% diag(sigma_u)

```

<!-- Suppose the correlation between the outcomes is 0.5. -->

<!-- ```{r Correlation} -->
<!-- corr = matrix(c(1,0.5,0.5,1),2,2) -->
<!-- ``` -->

Let's get the treatment mean curve, the standard deviations are assumed same as the placebo group.

```{r Treatment}
alpha=1
beta=1
effect_size_1 = 2.21*alpha/6
effect_size_2 = 5.38*beta/6

ym = xm + t(cbind(c(0,(rep(effect_size_1,6))), c(0,(rep(effect_size_2,6)))))

```

Finally Let's simulate the data.
```{r Generation}
X_placebo = array(NA, dim = c(nx, 2, 7))
X_dose_1 = array(NA, dim = c(ny, 2, 7))

for(i in 1:nx){
  X_placebo[i,,] = gen.fun.dep(xm, xs, corr, sigma)
}

for(i in 1:ny){
  X_dose_1[i,,] = gen.fun.dep(ym, xs, corr, sigma)
}

X_c = X_placebo
Y_c = X_dose_1
dim(X_c)
dim(Y_c)
```



Let's perform LRST on the generated data.

```{r LRST}
lrst.2arm(X_c, Y_c)
```

Let's get the estimated power.
```{r Power}
estimated_Power(0.05, X_c, Y_c)
```

Let's conduct the Multi-Arm LRST.

```{r Multi}
X_dose_2 = array(NA, dim = c(ny, 2, 7))
for(i in 1:ny){
  X_dose_2[i,,] = gen.fun.dep(ym, xs, corr, sigma)
}
X_c = X_placebo
Y = vector(length = 2, mode = "list")
Y[[1]] = X_dose_1
Y[[2]] = X_dose_2
lrst.MultiArm(X_c, Y)
```


# Citation

If you use this paper, please cite:

Xu, X., Ghosh, D., Luo, S., & CPP Integrated Parkinson’s Database. (2025). A novel longitudinal rank-sum test for multiple primary endpoints in clinical trials: Applications to neurodegenerative disorders. Statistics in Biopharmaceutical Research, (just-accepted), 1-17.

